Figures to insert in text

### 2018-04-04 kanerva-answer.tex
SDM Answer to Kanerva

\begin{theorem}
\emph{Given a reading address $x$ and a dimension $i$, the number of activated hard-locations with bitmatches at $i$ follows a normal distribution with }
\label{T2}
$\mathbf{E}[Z] = ph$ and $\mathbf{V}[Z] = hp(1-p) + p^2 H p_1 (1-p_1)$.

\end{theorem}

\begin{proof}
Let $Z$ be the number of activated hard locations with the same bit as the reading address:
$Z = \sum_{i=1}^{h} X_i$,

where $\mathbf{E}[h] = h$, $\mathbf{V}[h] = H p_1 (1-p_1)$, $p_1 = 2^{-n} \sum_{k=0}^{r} \binom{n}{k}$, and $X_i \sim \text{Bernoulli}(p)$.

By the central limit theorem, $Z$ is normally distributed.

Applying the law of total average and the law of total variance, $\mathbf{E}[Z] = \mathbf{E}[\mathbf{E}[Z | h]] = \mathbf{E}[ph] = p \mathbf{E}[h] = ph$, and $\mathbf{V}[Z] = \mathbf{E}[\mathbf{V}[Z|h]] + \mathbf{V}[\mathbf{E}[Z|h]] = \mathbf{E}[hp(1-p)] + \mathbf{V}[ph] = p(1-p) \mathbf{E}[h] + p^2 \mathbf{V}[h] = hp(1-p) + p^2 H p_1 (1-p_1)$.

@@

Let $s$ be the number of bitstrings written into memory (in our case, $s=10,000$), $h$ be the average number of activated hard locations ($h=1,071.85$), $H$ be the number of hard locations ($H=1,000,000$), and $\text{addr}_i$ be the $i$th bit of the hard location's address.

Let $\theta = \frac{sh}{H}$ be the average number of bitstrings written in each hard location, and $X_k \sim \text{Bernoulli}(p)$ (where $p = P(x_i = y_i | d(x, y) \le r)$). Thus,

\begin{align}
Y_i = \sum_{k=1}^{\theta} X_k \sim \mathcal{N}(\mu_1, \sigma_1^2)
\end{align}

Applying the law of total expectation, $\mathbf{E}[Y] = \mathbf{E}[\mathbf{E}[Y|\theta]] = \mathbf{E}[p \theta] = p \mathbf{E}[\theta] = p \theta$, since $\mathbf{E}[\theta] = \theta$.

Applying the law of total variance, $\mathbf{V}[Y] = \mathbf{E}[\mathbf{V}[Y|\theta]] + \mathbf{V}[\mathbf{E}[Y|\theta]] = \mathbf{E}[\theta p (1-p)] + \mathbf{V}[p \theta] = p(1-p) \mathbf{E}[\theta] + p^2 \mathbf{V}[\theta] = p(1-p) \theta + p^2 s^2 p_1 (1-p_1)/H$, since $\mathbf{V}[\theta] = s^2 p_1 (1-p_1)/H$.

Thus,

\begin{align}
\mu_1 &= p \theta \\
\sigma_1^2 &= p^2 s^2 p_1 (1 - p_1) / H
\end{align}

During a write operation, the counters are incremented for every bit 1 and decremented for every bit 0. So, after $s$ writes, there will be $\theta$ bitstrings written in each hard location, $Y_i$ bits 1, and $\theta - Y_i$ bits 0. Thus, $[\text{cnt}_i | \text{addr}_i = 1] = (Y_i) - (\theta - Y_i) = 2Y_i - \theta$; and $[\text{cnt}_i | \text{addr}_i = 0] = \theta - 2Y_i$.

Hence, as $\text{cnt}_i = 2Y_i - \theta$, $\mathbf{E}[2Y_i - \theta] = 2 \mathbf{E}[Y_i] - \theta$, and $\mathbf{V}[2Y_i - \theta] = 4 \mathbf{V}[Y_i]$, then,

\begin{align}
\left[ \text{cnt}_i | \text{addr}_i=1 \right] &\sim \mathcal{N}(\mu_2 = (2p-1) \theta, \sigma_2^2 = 4 \sigma_1^2)\label{cntaddr1} \\
\left[ \text{cnt}_i | \text{addr}_i=0 \right] &\sim \mathcal{N}(\mu_2 = -(2p-1) \theta, \sigma_2^2 = 4 \sigma_1^2)\label{cntaddr0}
\end{align}

In our case, $p=0.5529$, $s=10,000$, $h=1,071.85$, and $H=1,000,000$, so $\theta = 10.7185$ and $\text{cnt}_i \sim \mathcal{N}(\mu=1.1341, \sigma^2 = 10.7294)$. For ``random at x'', $p=0.5$, so $\mu = 0$ and $\sigma^2 = 10.8255$. See Figure \ref{fig:sdm-corr-counters}.

@@

For ``random written at x'', $p=0.5$ implies $\mu_2 = 0$, which implies $P(\text{cnt}_i > 0 | \text{addr}_i = 1) = P(\text{cnt}_i < 0 | \text{addr}_i = 0) = 0.5$, independently of the parameters because they will only affect the variance and the normal distribution is symmetrical around the average.

However, for ``x written at x'', $p=0.5529$ and the probabilities depend on $s$. For $s=10,000$, they are equal to 0.6354. For $s=20,000$, they are equal to 0.6867. For $s=30,000$, they are equal to 0.7232. The more random bitstrings are written into the memory, the more the hard locations point to themselves. See Figure \ref{fig:sdm-corr-prob} --- and notice that I still have to figure out why the mean is correct, but the standard deviation is not. As each of the $n$ counters of a hard location may be equal or not with the same probability, I assumed it would follow a Binomial distribution (and it worked for ``random at x'').

>>>>

Let $h$ be the number of activated hard locations. As the probability of activating a specific hard location is constant, $h \sim \text{Binomial}(H, p_1)$. Thus, $\mathbf{E}[h] = \mu_h = Hp_1$ and $\mathbf{V}[h] = \sigma^2_h = Hp_1(1-p_1)$, where $p_1 = 2^{-n} \sum_{k=0}^{r} \binom{n}{k}$.

Let $Z$ be the number of activated hard locations with the same bit as the reading address. Then, $Z = \sum_{i=1}^{h} X_i$, where $X_i \sim \text{Bernoulli}(p)$, where $p = P(x_i = y_i | d(x, y) \le r)$.

\begin{theorem}
Given a reading address $x$ and a dimension $i$, the number of activated hard-locations with bitmatches at $i$ follows a normal distribution with $\mathbf{E}[Z] = \mu_Z = p \mu_h$ and $\mathbf{V}[Z] = \sigma_Z^2 = p(1-p) \mu_h + p^2 \sigma^2_h$.
\label{T2}
\end{theorem}

\begin{proof}
As $P(973 < h < 1170) = 0.997$, by the central limit theorem, $Z$ may be approximated by a normal distribution.

Applying the law of total average, $\mathbf{E}[Z] = \mathbf{E}[\mathbf{E}[Z | h]] = \mathbf{E}[ph] = p \mathbf{E}[h] = p \mu_h$.

Applying the law of total variance, $\mathbf{V}[Z] = \mathbf{E}[\mathbf{V}[Z|h]] + \mathbf{V}[\mathbf{E}[Z|h]] = \mathbf{E}[hp(1-p)] + \mathbf{V}[ph] = p(1-p) \mathbf{E}[h] + p^2 \mathbf{V}[h] = p(1-p)\mu_h + p^2 \sigma^2_h$.
\end{proof}

@@

Let $s$ be the number of bitstrings written into memory (in our case, $s=10,000$) and $\text{addr}_i$ be the $i$th bit of the hard location's address.

Let $\theta$ be the average number of bitstrings written in each hard location. As there are $s$ bitstrings written into the memory, and the probability of activating a specific hard location is constant, $\theta \sim \text{Binomial}(s, p_1)$. Thus, $\mathbf{E}[\theta] = \mu_\theta = s p_1$ and $\mathbf{V}[\theta] = \sigma^2_\theta = s p_1 (1 - p_1)$.

Let $Y_i$ be the number of bitmatches in the $i$bit of a hard location's address after $s$ written bitstrings. Then, $Y_i = \sum_{k=1}^{\theta} X_k$.

\begin{theorem}
Giving the number of written bitstrings $s$, $\mathbf{E}[Y_i] = \mu_Y = p \mu_\theta$ and $\mathbf{V}[Y_i] = \sigma^2_Y = p(1-p) \mu_\theta + p^2 \sigma^2_\theta$.
\end{theorem}
\begin{proof}
Applying the law of total expectation, $\mathbf{E}[Y] = \mathbf{E}[\mathbf{E}[Y|\theta]] = \mathbf{E}[p \theta] = p \mathbf{E}[\theta] = p \mu_\theta$.

Applying the law of total variance, $\mathbf{V}[Y] = \mathbf{E}[\mathbf{V}[Y|\theta]] + \mathbf{V}[\mathbf{E}[Y|\theta]] = \mathbf{E}[\theta p (1-p)] + \mathbf{V}[p \theta] = p(1-p) \mathbf{E}[\theta] + p^2 \mathbf{V}[\theta] = p(1-p) \mu_\theta + p^2 \sigma^2_\theta$.
\end{proof}

During a write operation, the counters are incremented for every bit 1 and decremented for every bit 0. So, after $s$ writes, there will be $\theta$ bitstrings written in each hard location with $Y_i$ bitmatches and $\theta - Y_i$ non-bitmatches. Thus, $[\text{cnt}_i | \text{addr}_i = 1] = (Y_i) - (\theta - Y_i) = 2Y_i - \theta$ and $[\text{cnt}_i | \text{addr}_i = 0] = \theta - 2Y_i$.

\begin{theorem}
$\mathbf{E}[\text{cnt}_i | \text{addr}_i = 1] = \mu_{\text{cnt}} = (2p-1) \mu_\theta$ and $\mathbf{V}[\text{cnt}_i | \text{addr}_i = 1] = \sigma^2_\text{cnt} = 4p(1-p) \mu_\theta + (2p-1)^2 \sigma^2_\theta$.
\end{theorem}

\begin{proof}

$\mathbf{E}[\text{cnt}_i | \text{addr}_i = 1] = \mathbf{E}[2Y_i - \theta] = \mathbf{E}[2Y_i] - \mathbf{E}[\theta] = 2 \mathbf{E}[Y_i] - \mu_\theta = 2 p \mu_\theta - \mu_\theta = (2p-1) \mu_\theta$.

Applying the law of total variance, $\mathbf{V}[\text{cnt}_i | \text{addr}_i = 1] &= \mathbf{V}[2Y_i - \theta] = \mathbf{E}[\mathbf{V}[2Y_i - \theta | \theta]] + \mathbf{V}[\mathbf{E}[2Y_i - \theta | \theta]]$.

Let's solve each part independently. Thus,

$\mathbf{V}[2Y_i - \theta | \theta] = \mathbf{V}[2Y_i | \theta] = 4 \mathbf{V}[Y_i | \theta] = 4 \mathbf{V}[\sum_{k=1}^\theta X_k] = 4 \theta p (1-p)$.

$\mathbf{E}[\mathbf{V}[2Y_i - \theta | \theta]] = \mathbf{E}[4 \theta p (1-p)] = 4p(1-p) \mathbf{E}[\theta] = 4p(1-p) \mu_\theta$.

$\mathbf{E}[2Y_i - \theta | \theta] = 2 \mathbf{E}[Y_i | \theta] - \mathbf{E}[\theta | \theta] = 2p \theta - \theta = (2p-1) \theta$.

$\mathbf{V}[\mathbf{E}[2Y_i - \theta | \theta]] = \mathbf{V}[(2p-1) \theta] = (2p-1)^2 \mathbf{V}[\theta] = (2p-1)^2 \sigma^2_\theta$.
\end{proof}

\begin{theorem}
$\mathbf{E}[\text{cnt}_i | \text{addr}_i = 0] = - \mu_{\text{cnt}}$ and $\mathbf{V}[\text{cnt}_i | \text{addr}_i = 1] = \sigma^2_\text{cnt}$.
\end{theorem}
\begin{proof}
Notice that $[\text{cnt}_i | \text{addr}_i = 0] = -[\text{cnt}_i | \text{addr}_i = 1]$. Thus, $\mathbf{E}[\text{cnt}_i | \text{addr}_i = 0] = -\mathbf{E}[\text{cnt}_i | \text{addr}_i = 1]$ and $\mathbf{V}[\text{cnt}_i | \text{addr}_i = 0] = \mathbf{V}[\text{cnt}_i | \text{addr}_i = 1]$.
\end{proof}

In summary,

\begin{align}
\left[ \text{cnt}_i | \text{addr}_i=1 \right] &\sim \mathcal{N}(\mu_\text{cnt}, \sigma^2_\text{cnt})\label{cntaddr1} \\
\left[ \text{cnt}_i | \text{addr}_i=0 \right] &\sim \mathcal{N}(-\mu_\text{cnt}, \sigma^2_\text{cnt})\label{cntaddr0}
\end{align}

In our case, $p=0.5529$, $s=10,000$, and $H=1,000,000$, so $[\text{cnt}_i | \text{addr}_i=1] \sim \mathcal{N}(\mu=1.1341, \sigma^2 = 10.7184)$. For ``random at x'', $p=0.5$, so $\mu = 0$ and $\sigma^2 = 10.7185$. See Figure \ref{fig:sdm-corr-counters}.

@@

For ``random written at x'', $p=0.5$ implies $\mu_\text{cnt} = 0$, which implies $P(\text{cnt}_i > 0 | \text{addr}_i = 1) = P(\text{cnt}_i < 0 | \text{addr}_i = 0) = 0.5$, independently of the parameters because they will only affect the variance and the normal distribution is symmetrical around the average.

However, for ``x written at x'', $p=0.5529$ and the probabilities depend on $s$. For $s=10,000$, they are equal to 0.6354. For $s=20,000$, they are equal to 0.6867. For $s=30,000$, they are equal to 0.7232. The more random bitstrings are written into the memory, the more the hard locations point to themselves.

See Figure \ref{fig:sdm-corr-prob} and notice that I still have to figure out why the mean is correct, but the standard deviation is not. As each of the $n$ counters of a hard location may be equal or not with the same probability, I assumed it would follow a Binomial distribution (and it worked for ``random at x'').
